Word Learning as Network Growth: A Cross-linguistic Analysis

### 2018-11-12 journal/ms.Rmd
final journal draft

 Children tend to produce words earlier when they are connected to a variety of other words along both the phonological and semantic dimensions. Though this connectivity effect has been extensively documented, little is known about the underlying developmental mechanism.  One view suggests that learning is primarily driven by a network growth model where highly connected words in the child's early lexicon attract similar words. Another view suggests that learning is driven  by highly connected words in the external learning environment, instead of highly connected words in the early internal lexicon. The present study tests both scenarios systematically in both the phonological and semantic domains across 10 languages. We show that external connectivity in the learning environment drives growth in both production- and comprehension-based vocabularies, even controlling for word frequency and length. This pattern of findings suggests a word learning mechanism where children harness their statistical learning abilities to (indirectly) detect and learn highly connected words in the learning environment.

@@

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analyzing dense corpora capturing language learning in context [e.g., @roy2015]. One strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn. For example, within a lexical category, words that are more frequent in child-directed speech are acquired earlier [@goodman2008]. Other factors include word length, the mean length of utterances in which the word occurs, and concreteness [see @braginsky2016].

Besides these word-level properties, the structure of the lexicon (that is, how words relate to each other) also predicts the age of acquisition of words.  The lexical structure can be characterized in terms of a network where each node represents a word in the vocabulary, and each link between two nodes represents a relationship between the corresponding pair of words [e.g., @collins1975]. Previous studies have investigated early vocabulary structure by constructing networks using  a variety of word-word relations including shared semantic features, target-cue relationships in free association norms, co-occurrence in child directed speech, and phonological relatedness. These studies have found that children tend to produce words that have higher neighborhood density (i.e., high connectivity in the network) earlier, both at the phonological and the semantic level [@storkel2009; @hills2009; @hills2010; @stella2017; @carlson2014].

@@

These two proposals represent two divergent ideas about the role of lexical networks in acquisition. On the INT proposal, learning is driven by known words with high connectivity to other known words (Figure 1, left). Thus, the network structure is a causal factor in early word learning, that is, children might be relying on the organization of their past knowledge to determine future learning. For example, having a rich and organized knowledge about the domain of dinosaurs would facilitate the acquisition of new dinosaur-related words [@chi1983]. In contrast, on the EXT approach, learning is driven by the connectivity of words that are not known yet (Figure 1, right). Thus, the relevant network structure is not internally represented by children, and the related connectivity effect might be an epiphenomenon of some low level properties of the linguistic input. For example, highly connected words in the input could be more easily learned because of their contextual diversity [e.g., @smith2008] or because caregivers emphasize such words in child-directed speech [@macwhinney2014].

@@

@hills2009 investigated the growth of lexico-semantic networks in toddlers and found that growth did not proceed according to INT as was originally hypothesised by @steyvers2005, but rather according to EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.} This is a profound finding because it suggests that learning in the early stages is mostly driven by properties of the external input, regardless of how past knowledge is organized. However, this work has been done in the special case of networks that are based on 1) semantic associations, 2) production-based vocabularies, and 2) data from English-learning children, only. The extent to which this result depends on the domain (e.g., semantic vs. phonological connectivity), the vocabulary measure (production vs. comprehension) and culture/language is still unclear. In this work, we test the generality of the finding along these three dimensions.

First, we study the phonological network in addition to the semantic network. These two networks represent different ways the mental lexicon is structured. In particular, words that are neighbors in the semantic network (e.g., "cat", "dog") are not necessarily neighbors in the phonological network, and vice versa. Does the phonological network also predict word learning? Previous work did find an effect of words' connectivity in the phonological network on their age of learning [@storkel2009; @stella2017; @carlson2014]. In other words, words learned earlier in life tend to sound similar to many other words than a word learned later in life. However, this finding is *a priori* compatible with both INT and EXT, and previous studies did not explicitly compare these two mechanisms.  Here, we investigate whether phonological networks, like semantic networks, grow through EXT, or if they rather grow via INT (Figure\ \@ref(fig:growth)).
<!--
The phonological strcuture influence word processing in adults (Vitevitch, 2008;  Chan & Vitevitch, 2009; Chan & Vitevitch,2010; Vitevitch, Chan, and Roodenrys, 2012; Castro et al. 2018). Here we test whether the mechanism of networtk growth in the phonological domains mirrors the mechanism found in the semantic domains. -->

@@

Finally, we use developmental data in 10 languages. Lexical networks can show more or less cross-linguistic variability along both the semantic and phonological domains [@youn2016; @arbesman2010]. Besides, cultures might differ in the way caregivers talk to children [@kuhl1997; @cristia2017], and this difference in the input could influence the children's learning strategy. Thus, Cross-linguistic comparison is crucial to test what mechanism is cognitively universal and is used by all children, and what mechanism is specific to some patterns of learning that emerge due to the particulars of a given language or culture [@bates1987; @slobin2014]. 
 <!--
\footnote{The difference is more obvious in the phonological case where the network structure––and the distribution of word connectivity––can a priori change from language to language depending on various languistic conventions.}
-->

We test the growth scenarios using parent reports of children vocabularies and their normative age of acquisition [@fenson94].  Children may vary in their individual learning trajectories, but the aggregate data leads to an average learning pattern which is highly consistent. It is this normative trajectory that we model in the current study, following the steps of previous research using similar datasets [@storkel2009; @stella2017; @hills2009; @hills2010; @goodman2008; @braginsky2016]

The paper is organized as follows. First, we describe the datasets we used and explain how we constructed the networks. Second, we analyze static properties of words' connectivity in these networks (correlation with AoA and shape of the distribution) and we explain how these properties infom hypotheses about network growth. Next, we explicitly fit the two hypothesized growth mechanisms to the data. We investigate the extent to which the results obtained in @hills2009 generalize to phonological networks and comprehension-based vocabularies, and whether this generalization holds cross-linguistically.

@@

We used data from Wordbank [@frank2017], an open repository aggregating cross-linguistic language developmental data of the MacArthur-Bates Communicative Development Inventory (CDI), a parent report vocabulary checklist. Parent report is a reliable and valid measure of children's vocabulary that allows for the cost-effective collection of datasets large enough to test network-based models of acquisition [@fenson94]. When filling out a CDI form, caregivers are either invited to indicate whether their child "understands" (comprehension) or "understands and says" (production) each of about 400-700 words. For younger children (e.g., 8 to 18 months in the English data), both comprehension and production are queried, whereas for older children (16 to 36 months) only production is queried. Thus, we use data from younger children to test comprehension and data from older children to test production across 10 languages. Following previous studies [@storkel2009; @hills2009], we restrict our analysis to nouns. Table \@ref(tab:stats) gives an overview of the data.

@@

kable(statistics, format = "latex", escape = FALSE, booktabs = TRUE,
      linesep = "", format.args = list(big.mark = ","),
      caption = "Statistics for dataset we used.",
      col.names = c("Language",  "Nouns", "Ages", "Nouns", "Ages")) %>%
  add_header_above(c("", "Comprehension" = 2, "Production" = 2)) %>%
  column_spec(1, bold = TRUE) %>%
  kable_styling(position = "center")

@@

Since  the free association norms are available only in English, we used the hand-checked translation equivalents available in Wordbank, which allowed us to use the English association norms across languages. Using the same association data across languages does not necessarily mean that the resulting networks will be the same across languages, or that these networks will grow similarly. Indeed, though this approximation assumes that the semantic similarity measure is universal––which is a reasonable assumption [e.g., @youn2016], the set of words acquired by children as well as the timeline of this acquisition can still vary from language to language leading to possibility different learning strategies.

@@

In previous studies, two nodes were linked if they had an edit distance of 1 [@storkel2009; @carlson2014; @stella2017]. However, these studies reported a contribution of phonological networks to word learning when these networks were built using a rich adult vocabulary. However, since the focus of the current study is on the mechanism of growth, the networks should be based on the children's early vocabulary which, nevertheless, contains very few word pairs with an edit distance of 1. Thus, we increased the threshold from 1 to 2, that is, two nodes were related if their edit distance was equal to 1 or 2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker that those done with a threshold of 2.} The connectivity of a given node was characterized with its *degree*: the number of links it shares with other words.

@@

We start by analyzing word connectivity in the global (static) network. We constructed this network using nouns learned by the oldest age for which we have CDI data (e.g., in English this corresponds, in comprehension, to the network by 18 months, and in production, to the network by 30 months). This global network is the end-state towards which both INT and EXT converge by the last month of learning. Moreover, following @hills2009, we used this end-state network as a proxy for the external connectivity in the learning environment. Below we analyze properties of this global networks that are relevant to INT and/or EXT.  In order to compare various predictors on the same data, we restrict the analysis to the subset of nouns for which we had both semantic and phonological information in each language.

@@

Connectivity in the global network is directly related to EXT as it represents the explicit criterion this growth scenario uses to determine what words should be learned first (Figure \@ref(fig:growth)). Therefore, a direct consequence of an EXT-like growth scenario is a correlation between connectivity in the global network and the age of acquisition.\footnote{This correlation is also compatible with INT, although the causality is reversed. Indeed, from the perspective of this growth scenario, higher connectivity in the global network is caused by earlier learning, not the other way around. Some words end up being highly connected in the global network precisely because they happen to be acquired earlier and, therefore, have a higher chance of accumulating more links over time.} Figure \@ref(fig:corrAll) shows how the age of production and comprehension for each word varies as a function of its degree (or indegree for the semantic networks) as well as the correlation values. For ease of visual comparison, the predictor (i.e., the degree) was centered and scaled across languages.

The plots show, overall, a negative correlation between the month of acquisition and the degree.  In production data, the average correlation across languages was `r ps_ave` (SD=`r ps_sd`) for the semantic networks and `r pp_ave` (SD=`r pp_sd`) for the phonological networks.  In comprehension data, the average correlation was `r cs_ave` (SD=`r cs_sd`) for the semantic networks and `r cp_ave` (SD=`r cp_sd`) for the phonological networks. These results indicate that nouns with higher degrees are generally learned earlier, thus replicating previous findings in English (e.g., Storkel 2004, 2009; Hills et al. 2009) and extending these findings to ten different languages, generally, in both production- and comprehension-based vocabularies.

@@

We also analyzed the global network's degree distribution. The shape of this distribution is particularly relevant to INT as this growth scenario is known to generate networks with a power-law degree distribution, i.e., a distribution of the form $p(k) \propto \frac{1}{k^{\alpha}}$ [@barabasi99]. If the network displays this property, this fact would suggest, but not prove, an INT-like generative process. If the degree distribution does not follow a power law, this  would largely weaken the case for INT. The log-log plots are shown in Figure \@ref(fig:degreeDist). We fit a power law to each empirical degree distribution following the procedure outlined in @clauset09 and using a related R package [poweRlaw, @gillespie15].

@@

Overall, we could not reject the null hypothesis of a power-law distribution: The $p$-value was generally above 0.1 in almost all languages for both production and comprehension. That said, phonological networks had relatively larger cut-offs than semantic networks. As was suggested by @arbesman2010, these "truncated" power-laws in phonological networks–as well as the observed cross-linguistic variability in the value of the cut-offs–may reflect the various constraints that exist on word formation such as the number of phonemes in the language, the phonotactics (i.e., the way sound sequences are arranged in words), and the length of words. Such constraints may limit the number of words that are phonologically similar, thus leading to distributions which decay faster than a non-truncated power law.

In sum, the static properties of the global network are *a priori* compatible with both INT and EXT. In order to decide between these two developmental scenarios, we need to fit explicit growth models to the data.

@@

First, the results replicate Hills et al.'s original finding regarding the semantic network in English and production data, which is that this network grows by EXT, not by INT. Second, our results show that, generally speaking, this finding generalizes to comprehension, and holds across languages. This generalization was obtained in both the semantic\footnote{One could imagine that the fact of using English free association norms cross-linguistically would decrease the effect of non-English semantic networks because of possible cultural differences. However, our findings do not support this assumption, rather it supports our initial approximation about the universality of the semantic similarity measure.} and phonological domains.

@@

As for the factors of interest, i.e., semantic and phonological connectivity, we found cross-linguistic differences. Connectivity contributes to learning in some languages but not in other. In particular, semantic connectivity does not explain variance in English data beyond that explained by phonological connectivity, frequency and length. This contrasts with the original finding in @hills2009. However, this might be due to our using of a slightly different model (which included word length as a covariate) and a larger dataset. That said, and despite these apparent cross-linguistic differences, both phonological and semantic connectivity are significant predictors in the combined model.

@@

One possibility is that these patterns emerge from children’s use of statistical learning abilities [@saffran1996; @aslin2012; @smith2008]. The term "statistical learning" has been used in the developmental literature to describes the process by which one acquires information about their environment through tracking the frequency distribution of some elements (e.g., words) in different contexts. An important property of this kind of learning is that it occurs without explicit instructions and through mere exposure to the input. Previous work in the line of research has documented mechanisms which can explain the patterns found in the current study.

For example, in the semantic domain, growth according to EXT can be explained by a mechanism similar to cross-situational learning [@smith2008; @pinker2013; @yurovsky2015]. According to this mechanism, children track the co-occurrence of concrete nouns with their possible semantic referents. The referent of a word heard in only one naming situation can be ambiguous (e.g., when the word "ball" is heard for the first time in the presence of both a ball and a chair), but hearing the same word in a diversity of semantic contexts allows  the learner  to  narrow down the set of possible word-object mappings.  In our case, free association is related to contextual co-occurrence [@griffiths07], meaning that highly connected words will tend to occur in a variety of speech and referential contexts.  This fact makes such words easier to learn because they have more referential disambiguating cues across learning contexts, and crucially, even without knowing the entire set of words with witch they co-occur (hence the similarity with EXT). This possibility is supported by the finding that words' diversity of occurrence in child directed speech predicts their age of learning [@hills2010].

In the phonological case, network growth according to EXT is also compatible with a scenario whereby children are tracking lwo level statistical patterns, e.g., high probability sound sequences. Indeed, connectivity in the phonological network is inherently correlated with phonotactic probability [@vitevitch1999]. That is, highly connected words tend to be made of frequent sound sequences. Children are sensitive to local phonotactic regularities [@jusczyk1994] and this sensitivity might lead them to learn higher-probability words more easily [@storkel2001]. This explanation is supported by computational simulations that shows how learning general phonotactics patterns create ``well-worn paths'' which allow the models to represent several distinct but phonologically neighboring words [@dell1993; @takac2017].

@@

This work shares a number of limitations with previous studies using similar research strategy and datasets. Chief among these limitations is the fact that the age of word acquisition is computed using different children at different ages (due to the fact that available CDI data is mainly cross-sectional). Although this measure has proven highly consistent [@fenson94], it led us to focus on studying the learning mechanism of the "average" child. Individual trajectories, however, could show different learning patterns. For example, using longitudinal data @beckage2011 found differences between typical and late talkers in terms of the semantic network structure. Besides, although our study endorses the externally-driven account of network growth, this does not mean individual children never use some variant of INT or some combination of both INT and EXT (Beckage and Colunga, under review). For example, some children develop "islands of expertise", that is, well organized knowledge about a certain topic (e.g., birds or dinosaurs). This prior knowledge enables these children to learn new related words more easily [e.g., @chi1983].

>>>>

 Children tend to produce words earlier when they are connected to a variety of other words along both the phonological and semantic dimensions. Though this connectivity effect has been extensively documented, little is known about the underlying developmental mechanism.  One view suggests that learning is primarily driven by a network growth model where highly connected words in the child's early lexicon attract similar words. Another view suggests that learning is driven  by highly connected words in the external learning environment, instead of highly connected words in the early internal lexicon. The present study tests both scenarios systematically in both the phonological and semantic domains across 10 languages. We show that external connectivity in the learning environment drives growth in both production- and comprehension-based vocabularies, even controlling for word frequency and length. This pattern of findings suggests a word learning mechanism where children harness their statistical learning abilities to detect and learn highly connected words in the learning environment.

@@

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analyzing dense corpora capturing language learning in context [e.g., @roy2015]. One strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn. For example, within a lexical category, words that are more frequent in child-directed speech are acquired earlier [@goodman2008]. Other factors include word length and the mean length of utterances in which the word occurs [e.g., @braginsky2016; @swingley2018].

Besides these word-level properties, the structure of the lexicon (that is, how words relate to one another) also predicts the Age of Acquisition (AoA) of words.  The lexical structure can be characterized in terms of a network where each node represents a word in the vocabulary, and each link between two nodes represents a relationship between the corresponding pair of words [e.g., @collins1975]. Previous studies have investigated early vocabulary structure by constructing networks using  a variety of word-word relations including shared semantic features [@mcrae2005], target-cue relationships in free association norms [@nelson1998], co-occurrence in child directed speech [@macwhinney2014], and phonological relatedness [@vitevitch2008]. These studies have generally found that children tend to produce words that have higher neighborhood density (i.e., high connectivity in the network) earlier, both at the phonological and the semantic level [@storkel2009; @hills2009; @hills2010; @stella2017; @carlson2014].

@@

These two proposals represent two divergent ideas about the role of lexical networks in acquisition. On the INT proposal, learning is driven by known words with high connectivity to other known words (Figure 1, left). Thus, the network structure is a causal factor in early word learning, that is, children rely on the organization of their past knowledge to determine future learning. For example, having a rich and organized knowledge about the domain of dinosaurs would facilitate the acquisition of new dinosaur-related words [@chi1983]. In contrast, on the EXT approach, learning is driven by the connectivity of words that are not known yet (Figure 1, right). Thus, the relevant network structure is not internally represented by children, and the observed connectivity effect might be an epiphenomenon of some properties of the linguistic input. For example, highly connected words in the input could be more easily learned because of their contextual diversity [e.g., @smith2008] or because caregivers emphasize such words in child-directed speech [@macwhinney2014].

@@

@hills2009 investigated the growth of lexico-semantic networks in toddlers and found that growth did not proceed according to INT as was originally hypothesised by @steyvers2005, but rather according to EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.} This is a profound finding because it suggests that learning in the early stages is mostly driven by properties of the external input, regardless of how past knowledge is organized. However, this work has been done in the special case of networks that are based on 1) semantic associations, 2) production-based vocabularies, and 3) data from English-learning children, only. The extent to which this result depends on the domain (e.g., semantic vs. phonological connectivity), the vocabulary measure (production vs. comprehension) and culture/language is still unclear. In this work, we test the generality of the finding along these three dimensions.

First, we study the phonological network in addition to the semantic network. These two networks represent different ways the mental lexicon is structured [@beckage2016]. In particular, words that are neighbors in the semantic network (e.g., "cat", "dog") are not necessarily neighbors in the phonological network, and vice versa. Does the phonological network also predict word learning? Previous work did find an effect of words' connectivity in the phonological network on their age of learning [@storkel2009; @stella2017; @carlson2014]. In other words, words learned earlier in life tend to sound similar to many other words than a word learned later in life. However, this finding is *a priori* compatible with both INT and EXT, and previous studies did not explicitly compare these two mechanisms.  Here, we investigate whether phonological networks, like semantic networks, grow through EXT, or if they rather grow via INT (Figure\ \@ref(fig:growth)).
<!--
The phonological strcuture influence word processing in adults (Vitevitch, 2008;  Chan & Vitevitch, 2009; Chan & Vitevitch,2010; Vitevitch, Chan, and Roodenrys, 2012; Castro et al. 2018). Here we test whether the mechanism of networtk growth in the phonological domains mirrors the mechanism found in the semantic domains. -->

@@

Finally, we use developmental data in 10 languages. Lexical networks can show more or less cross-linguistic variability along both the semantic and phonological domains [@youn2016; @arbesman2010]. Besides, cultures might differ in the way caregivers talk to children [@kuhl1997; @cristia2017], and this difference in the input could influence the way in which the children's networks grow. Thus, Cross-linguistic comparison is crucial to test what growth mechanism is cognitively universal and is used by all children, and what mechanism is specific to some patterns of learning that emerge due to the particulars of a given language or culture [@bates1987; @slobin2014]. 
 <!--
\footnote{The difference is more obvious in the phonological case where the network structure––and the distribution of word connectivity––can a priori change from language to language depending on various languistic conventions.}
-->

We test the growth scenarios using parent reports of children vocabularies and their normative age of acquisition [@fenson94].  Children may vary in their individual learning trajectories, but the aggregate data leads to an average learning pattern which is highly consistent. It is this normative trajectory that we model in the current study, following in this regard the steps of previous research using similar datasets [@storkel2009; @stella2017; @hills2009; @hills2010; @goodman2008; @braginsky2016]

The paper is organized as follows. First, we describe the datasets we used and explain how we constructed the networks. Second, we analyze static properties of words' connectivity in these networks (correlation with age of acquisition and shape of the distribution) and we explain how these properties infom hypotheses about network growth. Next, we explicitly fit the two hypothesized growth mechanisms to the data. We investigate the extent to which the results obtained in @hills2009 generalize to phonological networks and comprehension-based vocabularies, and whether this generalization holds cross-linguistically.

@@

We used data from Wordbank [@frank2017], an open repository aggregating cross-linguistic language developmental data of the MacArthur-Bates Communicative Development Inventory (CDI), a parent report vocabulary checklist. Parent report is a reliable and valid measure of children's vocabulary that allows for the cost-effective collection of datasets large enough to test network-based models of acquisition [@fenson94]. When filling out a CDI form, caregivers are either invited to indicate whether their child "understands" (comprehension) or "understands and says" (production) each of about 400-700 words. For younger children (e.g., 8 to 18 months in the English data), both comprehension and production are queried, whereas for older children (16 to 36 months) only production is queried. Due to these limitations, we use data from younger children to test comprehension and data from older children to test production. Following previous studies [@storkel2009; @hills2009], we restrict our analysis to nouns. Table \@ref(tab:stats) gives an overview of the data we used.

@@

kable(statistics, format = "latex", escape = FALSE, booktabs = TRUE,
      linesep = "", format.args = list(big.mark = ","),
      caption = "Statistics for the dataset we used.",
      col.names = c("Language",  "Nouns", "Ages", "Nouns", "Ages")) %>%
  add_header_above(c("", "Comprehension" = 2, "Production" = 2)) %>%
  column_spec(1, bold = TRUE) %>%
  kable_styling(position = "center")

@@

Since  the free association norms are available only in English, we used the hand-checked translation equivalents available in Wordbank, which allowed us to use the English association norms across 10 languages. Using the same association data across languages does not necessarily lead to similar networks. Indeed, though this approximation assumes that the semantic similarity measure is universal–which is a reasonable approximation [e.g., @youn2016], the set of words acquired by children as well as the timeline of this acquisition can still vary from language to language leading to possibility different learning strategies.

@@

In previous studies, two nodes were linked if they had an edit distance of 1 [@storkel2009; @carlson2014; @stella2017]. However, these studies reported a contribution of phonological connectivity to word learning when networks were built using a rich adult vocabulary. Since the focus of the current study is on the mechanism of growth, the networks should be based on the children's early vocabulary. The latter, nevertheless, contains very few word pairs with an edit distance of 1. Thus, we increased the threshold from 1 to 2, that is, two nodes were related if their edit distance was equal to 1 or 2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker that those obtained with a threshold of 2.} The connectivity of a given node was characterized with its *degree*: the number of links it shares with other words.

@@

We start by analyzing word connectivity in the global (static) network. We constructed this network using nouns learned by the oldest age for which we have CDI data (e.g., in English this corresponds, in comprehension, to the network by 18 months, and in production, to the network by 30 months). This global network is the end-state towards which both INT and EXT converge by the last month of learning. Moreover, following @hills2009, we used this end-state network as a proxy for the external connectivity in the learning environment. Below we analyze properties of this global networks that may a priori hint at an INT- or EXT-like growth.  In order to compare various predictors on the same data, we restrict the analysis to the subset of nouns for which we had both semantic and phonological information in each language.

@@

Connectivity in the global network is directly related to EXT as it represents the explicit criterion this growth scenario uses to determine what words should be learned first (Figure \@ref(fig:growth)). Therefore, a direct consequence of an EXT-like growth scenario is a correlation between connectivity in the global network and the age of acquisition. This correlation is also compatible with INT, although the causality is reversed: higher connectivity in the global network is caused by earlier learning, not the other way around. Some words end up being highly connected in the global network precisely because they happen to be acquired earlier and, therefore, have a higher chance of accumulating more links over time. Thus, the correlates between connectivity in the end-state network and AoA is an important blusprint of both EXT and INT. If there is no such correlation, neither growth scenarios can be posited as a learning mechanism.

Figure \@ref(fig:corrAll) shows how the age of production and comprehension for each word varies as a function of its degree (or indegree for the semantic networks) as well as the correlation values. For ease of visual comparison, the predictor (i.e., the degree) was centered and scaled across languages. The plots show, overall, a negative correlation between the month of acquisition and the degree.  In production data, the average correlation across languages was `r ps_ave` (SD=`r ps_sd`) for the semantic networks and `r pp_ave` (SD=`r pp_sd`) for the phonological networks.  In comprehension data, the average correlation was `r cs_ave` (SD=`r cs_sd`) for the semantic networks and `r cp_ave` (SD=`r cp_sd`) for the phonological networks. These results indicate that nouns with higher degrees are generally learned earlier, thus replicating previous findings in English (e.g., Storkel 2004, 2009; Hills et al. 2009) and extending these findings to 10 different languages, generally, in both production- and comprehension-based vocabularies.

@@

We also analyzed the global network's degree distribution. The shape of this distribution is particularly relevant to INT as this growth scenario is known to generate networks with a power-law degree distribution, i.e., a distribution of the form $p(k) \propto \frac{1}{k^{\alpha}}$ [@barabasi99]. If the end-state network displays this property, this fact would suggest, but not prove, an INT-like generative process. If, however, the degree distribution is very different from a power law, this would significantly weaken the case for INT. The log-log plots are shown in Figure \@ref(fig:degreeDist). We fit a power law to each empirical degree distribution following the procedure outlined in @clauset09 and using a related R package [poweRlaw, @gillespie15].

@@

Overall, we could not reject the null hypothesis of a power-law distribution: The $p$-value was generally above 0.1 in almost all languages for both production and comprehension. That said, phonological networks had relatively larger cut-offs than semantic networks. As was suggested by @arbesman2010, these "truncated" power-laws in phonological networks –as well as the cross-linguistic variability in the value of the cut-offs– may reflect the various constraints that exist on word formation such as the number of phonemes in the language, the phonotactics (i.e., the way sound sequences are arranged in words), and the length of words. Such constraints may limit the number of words that are phonologically similar, thus leading to distributions which decay faster than a non-truncated power law.

In sum, the static properties of the end-state network are *a priori* compatible with both INT and EXT. In order to decide between these two developmental scenarios, we need to fit explicit growth models to the data.

@@

First, the results replicate Hills et al.'s original finding regarding the semantic network in English and the production-based vocabulary, which is that this network grows by EXT, not by INT. Second, our results show that, generally speaking, this finding generalizes to comprehension-based vocabulary, and holds across languages. This generalization was obtained in both the semantic\footnote{One could imagine that the fact of using English free association norms cross-linguistically would decrease the effect of non-English semantic networks because of possible cultural differences. However, our findings do not support this assumption, rather it supports our initial approximation about the universality of the semantic similarity measure.} and phonological domains.

@@

As for the factors of interest, i.e., semantic and phonological connectivity, we found cross-linguistic differences. Connectivity contributes to learning in some languages but not in other. In particular, semantic connectivity does not explain variance in English data beyond that explained by phonological connectivity, frequency and length. This contrasts with the original finding in @hills2009. However, this might be due to our using a slightly different model (which included word length as a covariate) and a larger dataset. That said, and despite these apparent cross-linguistic differences, both phonological and semantic connectivity are significant predictors in the combined model (Figure \@ref(fig:staticAll)).

@@

One possibility is that these patterns emerge from children’s use of statistical learning abilities [@saffran1996; @aslin2012; @smith2008]. The term "statistical learning" has been used in the developmental literature to describes the process by which one acquires information about their environment through tracking the frequency distribution of some elements (e.g., words) in different contexts. An important property of this kind of learning is that it occurs without explicit instructions and through mere exposure to the input. Previous work in the line of research has documented specific mechanisms which can explain the patterns found in the current study.

For example, in the semantic domain, growth according to EXT can be explained by a mechanism similar to cross-situational learning [@smith2008; @pinker2013; @yurovsky2015]. According to this mechanism, children track the co-occurrence of concrete nouns with their possible semantic referents. The referent of a word heard in only one naming situation can be ambiguous (e.g., when the word "ball" is heard for the first time in the presence of both a ball and a chair), but hearing the same word in a diversity of semantic contexts allows the learner to narrow down the set of possible word-object mappings. In our case, free association (used to determine semamtnic network connectivity) is related to contextual co-occurrence [@griffiths07; @fourtassi2013], meaning that highly connected words will tend to occur in a variety of speech and referential contexts.  This fact makes such words easier to learn because they have more referential disambiguating cues across learning contexts. Crucially, children can learn these words without knowing the other words with witch they co-occur (hence the similarity with EXT). This possibility is supported by the finding that words' diversity of occurrence in child directed speech predicts their age of learning [@hills2010].

In the phonological case, network growth according to EXT is also compatible with a scenario whereby children are tracking lwo level statistical patterns, e.g., high probability sound sequences. Indeed, connectivity in the phonological network is inherently correlated with phonotactic probability [@vitevitch1999]. That is, highly connected words tend to be made of frequent sound sequences. Children are sensitive to local phonotactic regularities [@jusczyk1994] and this sensitivity might lead them to learn higher-probability words more easily [@storkel2001]. This explanation is supported by computational simulations that show how learning general phonotactics patterns create ``well-worn paths'' which allow the models to represent several distinct but phonologically neighboring words [@dell1993; @takac2017].

@@

This work shares a number of limitations with previous studies using similar research strategy and datasets. Chief among these limitations is the fact that the age of word acquisition is computed using different children at different ages (due to the fact that available CDI data is mainly cross-sectional). Although this measure has proven highly consistent [@fenson94], it led us to focus on studying the learning mechanism of the "average" child. Individual trajectories, however, could lead to different netwtok strcutures and show different learning patterns. For example, using longitudinal data @beckage2011 found differences between typical and late talkers in terms of the semantic network structure. Besides, although our study endorses the externally-driven account of network growth, this does not mean individual children never use some variant of INT or some combination of both INT and EXT (Beckage and Colunga, under review). For example, some children develop "islands of expertise", that is, well organized knowledge about a certain topic (e.g., birds or dinosaurs). This prior knowledge enables these children to learn new related words more easily [e.g., @chi1983].
